---
title: "Project3"
author: "Yuzhi Li and Annie Brinza"
date: "2022-11-10"
output: html_document
params:
  channel: "data_channel_is_lifestyle"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(corrplot)
```

# **Introduction**

# **Data**
The first step we need to take is reading in the data. Then we will need to 
subset to our channel of interest. In this case, that channel is `params$channel`.

```{r read_data}
df <- read_csv("data/OnlineNewsPopularity.csv", show_col_types = FALSE)
channel_df <- df %>% 
  filter(data_channel_is_lifestyle == 1) %>% 
  select(-starts_with("data_channel_is"))
```


# **Summarizations**
In this section, we will produce some basic yet meaningful summary statistics and
plots about the training data we're working with. We will focus most closely on how
our data relates to the shares variable.

First, we will look into title length. The length of a title can affect whether someone
is willing to share an article, since if the title gets cut off when sharing, a 
person might be less likely to share. This could differ by channel. In the summary
below, pay attention to the average title length and see how it differs from the
other channels. 

```{r summary_1}
summary(channel_df$n_tokens_title)
```

To continue this investigation, we will plot title length against shares. If there
is a negative downward trend, that means the longer the title length, the less likely
a person is to share.

```{r graph_1}

```


# **Modeling**

# **Comparison**